% ---- ETD Document Class and Useful Packages ---- %
\documentclass{ucetd}  % bib file is coded in this
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{multirow}
\usepackage{microtype}
\usepackage{amssymb}
\usepackage{docmute}
\usepackage{morefloats}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{afterpage}
\usepackage{longtable}
%\usepackage[nottoc,numbib]{tocbibind}

\newenvironment{section_abstract}{\rightskip1in}{}



%% Use these commands to set biographic information for the title page:
\title{Remodeling the Fossil Record: \protect\\Analysis of Emergent Evolutionary and Ecological Patterns}
\author{Peter David Smits}
\department{Committee on Evolutionary Biology}
\division{Biological Sciences \protect\\and the Pritzker School of Medicine}
\degree{Doctor of Philosophy}
\date{June 2017}

%% Use these commands to set a dedication and epigraph text
\dedication{Dedication Text}
\epigraph{Epigraph Text}



\begin{document}
%% Basic setup commands
% If you don't want a title page comment out the next line and uncomment the line after it:
\maketitle
%\omittitle

% These lines can be commented out to disable the copyright/dedication/epigraph pages
\makecopyright
\makededication
\makeepigraph


%% Make the various tables of contents
\tableofcontents
\listoffigures
\listoftables

% Enter Acknowledgements here
\acknowledgments
My academic journey started many years ago and has crossed multiple institutions and continents. 

First and foremost I would like to thank my advisors Kenneth D. Angielczyk and Michael Foote. Their support and advice helped guide my often erratic thinking into concrete products. Michael has been an incredible mentor and his constant pushing for me to ``speak in English'' has greatly improved my communication skills. Ken is an excellent collaborator and has been incredibly responsive even though we very rarely see each other in person; his help with crafting the narrative and improving the impact of my research has been an incredible boon. I look forward to continued collaboration with both of them.

My dissertation committee has been an invaluable resource for expanding my thinking and helping me translate my ideas for a broader audience than just myself. I am very lucky to have had such knowledgeable, questioning, and responsive group of readers. Thank you P. David Polly, Richard Ree, and Graham Slater for your incredible support. 

I've interacted with many professors along my academic journey, many of whom have been great influences on my career. 

I owe a big thanks to my undergraduate advisor Gregory Wilson. Greg took a risk on me and let me contribute directly to his research program, which was an incredible experience that led to my first paper. I learned a lot about fossil mammals, the Cretaceous, and field work from Greg and I owe what little ``real'' paleontology \textit{bona fides} to him and for that I'm forever grateful. 

My Master's advisor Alistair Evans taught me the importance of not overwhelming your audience when trying to explain the context of scientific study, something which I still work on everyday. Al also gave me an incredible degree of freedom to explore what questions interest me, something which really shaped my approach to research; I'm nothing if not independent. And while I may not have worked directly on mammal morphology for my dissertation, the dietary and functional perspective that pervades Alistair's research has never been far from my mind. Thank you.

There have been countless colleagues, professors, and peers who have helped shaped not only my dissertation but also my thinking in general and sadly I cannot name them all. Nancy Simmons, Jim Kenagy, Jeff Bradley, and Donald Grayson all provided me with research and/or discussion opportunities which have proved invaluable for my growth as a scientist. I will never forget Nancy telling me that the most important thing I could do to increase the impact of my research was to grow my statistical knowledge; this is a lesson I've truly taken to heart. Stewart Edie has been a fantastic collaborator and I look forward to our continued work together. My officemates Dallas Krentzel and Max Winston have been an excellent sounding board for questions, frustrations, and curiosities. The graduate student community of Darwinian Sciences and the paleontology section of Geophysical Sciences have also been a source of continued strength and support; I can't count the number of times TG has saved my mood from completely nose-diving. A big thank you to you all.

I could not have completed this dissertation without the incredible support of my friends and family. 

My parents, Alexander Smits and Louise Handelman, have been an unending source of support through out this entire process. They never questioned my goals or career decisions, and have provided such important life advice and unconditional love that I cannot even begin to express my appreciation and gratefulness for all the support you've given me over the years. I also want to give a special thanks to my brother, James Smits. I could not have asked for a more supportive and helpful brother. Thank you; I would never have been able to do this without all of you.

While I have moved all over the globe, my friends have always been an important source of distraction and happiness that I could not have dreamed surviving without. Thank you to Jessica Escott, Andrew Carson, Tyler Cooper, Clint Bland, Connor O'Donnell, Colin Kyle, Darcy Ross, Elizabeth Sander, Julie Szymaszek, Brian Waligorski, Brian Goodrich, anyone who has every played Dungeons and Dragons with me, and countless others. Without your presence I would never have had release.

Finally, I would like to thank my wonderful partner Megan Boatright who has been an incredible anchor from before my dissertation proposal all the way to my dissertation defense. Our lake-front conversations helped me coalesce around my research topics, our late-nights working together help me write both my proposal and my dissertation, and our many journeys together re-opened my eyes to the beauty and joy in the world. I love you Megan with all my heart.



\include{abstract}

\mainmatter
% Main body of text follows

% introductory stuff
\include{intro}

% chapter 1
\include{chapter_death_taxa/plain_survival}

% chapter 2
\include{chapter_preserve/plain_brachiopod}

% chapter 3
\include{chapter_coping/plain_coping}

% conclusion stuff
\include{conclusion}

% Format a LaTeX bibliography
\makebibliography

% Figures and tables, if you decide to leave them to the end
%\input{figure}
%\input{table}
\input{chapter_death_taxa/plain_mass_tab}

\end{document}
