\documentclass[12pt,letterpaper]{article}

\usepackage{amsmath, amsthm}
\usepackage{microtype, parskip}
\usepackage[comma,numbers,sort&compress]{natbib}
\usepackage{lineno}
\usepackage{docmute}
\usepackage{caption, subcaption, multirow, morefloats, rotating}
\usepackage{wrapfig}
\usepackage{attrib}

\frenchspacing

\begin{document}

\section*{Discussion}

Both the composition of a species pool and its environmental context change over time, though not necessarily at the same rate or concurrently. Local communities, whose species are drawn from the regional species pool, have ``roles'' in their communities defined by their interactions with a host of biotic and abiotic interactors (i.e. a species' niche). For higher level ecological characterizations like ecotypes and guilds, these roles are broad and not defined by specific interactions but by the genre of interactions species within that grouping participate in. The diversity of species within an ecotype or guild can be stable over millions of years despite constant species turnover \citep{Jernvall2004,Slater2015c,Valkenburgh1999}. This implies that the size and scope of the role of an ecotype or guild in local communities, and the regional species pool as a whole, is preserved even as the individual interactors change. This also implies that the structure of regional species pools can be constant over time despite a constantly changing set of ``players.'' There is even evidence that functional groups are at least partially self-organizing and truly emergent \citep{Scheffer2006a}.

Comparison of the results from the posterior predictive checks for the pure-presence and birth-death models supports the conclusion that regional species pool dynamics cannot simply be described by a single occurrence probability and are instead the result of the interplay between the origination and extinction processes. Additionally, changes to the ecotypic composition and diversification rate of the North American regional species pool are driven primarily by variation in origination and not extinction (Fig. \ref{fig:macro_values}). These aspects of how regional species pool diversity is shaped are not directly observable in studies of the Recent where time scales are short and macroevolutionary dynamics are inferable solely from phylogeny \citep{Fritz2013a,Price2016b}.

Extinction rate for the entire regional species pool through time is highly variable and demonstrates a saw-toothed pattern with no obvious temporal trends. While a constant mean extinction rate is consistent with previous observation \citep{Alroy1996a,Alroy2000g}, the degree to which mammal extinction rate is actually variable may not have been equally appreciated as it has been for the marine invertebrate record \citep{Foote2000,Foote2000a,Foote2006,Foote2010}. What is most consistent with previous observations, however, is that diversity seems to be most structured by changes to origination rather than changes to extinction \citep{Alroy1996a,Alroy2000g}.

Comparison of the ecotype specific diversity histories adds a considerable degree of nuance to broad narrative of shifts in functional composition of the North American mammal species pool as being gradual (Fig. \ref{fig:ecotype_diversity}). While most ecotypes do not experience sudden shifts in origination or extinction rate (Fig. \ref{fig:ecotype_birth}, \ref{fig:ecotype_death}). As with the diversification rate of the entire species pool, the diversification of individual ecotypes seem principally driven by origination and not extinction. Instead, while species seem to originate in waves (Fig. \ref{fig:ecotype_birth}), they appear to leave the regional species pool in an uncoordinated and individual manner (Fig. \ref{fig:ecotype_death}) which could be considered consistent with the maxim that all species respond differently to environmental change \citep{Blois2009}. Note, however, this result characterizes the entire North American mammal regional species pool and thus may not reflect the dynamics of individual local communities.

The few large-magnitude, but temporary, increases in ecotype-specific origination rate occur in digitigrade carnivores, digitigrade herbivores, plantigrade herbivores, and unguligrade hebivores. Importantly, the large peak in diversification and origination rates 16 Mya (Fig. \ref{fig:macro_values}) appears driven almost entirely by a massive increase in the origination rate of unguligrade herbivores (Fig. \ref{fig:ecotype_birth}). Additionally, there is some evidence that the origination probabilities of these ecotypes are correlated (Fig. \ref{fig:origin_corr}, \ref{fig:origin_corr_graph}). While this result does not mean that there are large and sudden cross-ecotype changes to the regional species pool, it does suggest that additions to the species pool do not occur in individual ecotypes idiosyncratically.

Arboreal taxa disappear from the regional species pool over the Cenozoic, with long term decline over the Paleogene leading to the disappearance by start of Neogene \(\sim\)22 Mya. This is partially consistent with one of the two possible patterns presented here and in \citet{Smits2015b} that would result in arboreal taxa having a greater extinction risk than other ecotypes: the Paleogene and Neogene were different selective regimes and, while the earliest Cenozoic may have been neutral with respect to arboreal taxa, they disappeared quickly over the Cenozoic which may account for their higher extinction risk. However, these result add some nuance to this scenario as arboreal taxa were declining throughout the Paleogene instead of maintaining a flat diversity as hypothesized \citep{Smits2015b}. I interpret the decline of arboreal taxa through out the Paleogene to mean that the shift from closed to open environments began in the Paleogene and led to increasingly hostile environments for arboreal taxa as opposed to being a sudden change in selective regime between the Paleogene and Neogene. In addition to all arboreal taxa, the diversity of plantigrade and scansorial insectivores decreases with time (Fig. \ref{fig:ecotype_diversity}).

Digitigrade carnivores have a relatively stable diversity history through the Cenozoic and can be characterized as varying around a constant mean diversity. This ecotype has a large amount of overlap with the carnivore guild which has been the focus of much research \citep{Slater2015c,Valkenburgh1999,Pires2015a,Janis1993c}. This result is consistent with some form of ``control'' on the diversity of this ecotype, such as diversity-dependent diversification \citep{Slater2015c,Silvestro2015b,Valkenburgh1999}.

Both digitigrade and unguligrade herbivores increase in diversity over the Cenozoic. The increase of these cursorial forms is consistent with the gradual opening up of the North American landscape \citep{Blois2009,Stromberg2005,Graham2011a} and may indicate a long-term shift in the interactors associated with those ecotypes leading to increased contribution to the regional species pool. This result may be comparable to the increasing percentage of hypsodont (high-crowned teeth) mammals in the Neogene of Europe being due to an enrichment of hyposodont taxa and not a depletion of non-hypsodont taxa. Smaller scale increases in fossorial herbivore species, and a lesser extent plantigrade herbivores, suggests that the increase of interactors may be associated mostly with the herbivore dietary category with locomotor category tempering that relationship. These results support the conclusion that the increase in digitigrade and unguligrade herbivores is the result of an enrichment of these ecotypes as opposed to being caused by the depletion of other herbivorous ecotypes; this is further supported by the lack of major changes to the number of extinctions of all herbivore ecotypes (Fig. \ref{fig:ecotype_death}).

An association between plant phase and differences in ecotype occurrence or origination-extinction probabilities is interpreted to mean that an ecotype enrichment or depletion is due to associations between that ecotype and whatever plants are dominant at that time. Plant phase clearly structures the occurrence and origination probability time series (Fig. \ref{fig:eco_occur}, \ref{fig:eco_origin}). These differences in occurrence or origination translate to the estimates of diversity and diversification rate; the largest spike in diversity, diversification rate, and origination rate all correspond to the onset of the last plant phase (Fig. \ref{fig:macro_values}). The clearest example of the diversity of an ecotype increasing at this particular transition is in scansorial carnivores (Fig. \ref{fig:ecotype_diversity}); similar shifts in other ecotypes are much more subtle, as was previously noted for fossorial insectivores. 

Interestingly, for all of the ecotypes with sudden changes in diversity at this transition the change is an increase, even if only temporarily. There are two interpretations of these results. A biological interpretation of this result is that, because plant phase associations are only with occurrence or origination probabilities and not survival, these ecotypes were well suited for the newly available mammal-plant interactions due to the increased modernization of their floral context \citep{Graham2011a}. Alternatively, the increase in diversity associated with the third plant phase may be caused by the edge effect in origination probability that is artificially inflating the number of origination events (Fig. \ref{fig:eco_origin}). However, the estimated number of origination events does not have a tremonedous spike at this transition, nor is a major increase in the number of origination events sustained (Fig. \ref{fig:ecotype_birth}).

There are fewer, less obvious shifts in diversity surrounding the transition from the first to second plant phase, with the following ecotypes having apparent shifts in diversity at 50 My: plantigrade carnivores (down), arboreal omnivores (down), and scansorial omnivores (down). Arboreal insectivore peak diversity also occurs 50 Mya, and is then followed by a steep decline in diversity till 30 Mya when this ecotype is lost from the species pool. Because plant phase has been found to structure occurrence/origination (Fig. \ref{fig:eco_occur}, \ref{fig:eco_origin}), but not survival (Fig. \ref{fig:eco_survival}), my interpretation of these results is that new species were not entering the system because there were fewer available mammal-plant interactions available for those ecotypes. Instead, these ecotypes were poorly suited for the newly available mammal-plant interactions brought upon by the changing environmental context \citep{Graham2011a}.

The temperature covariates are found to have similar effects on occurrence and origination probabilities (Tables \ref{tab:occur_temp}, \ref{tab:origin_temp}). Temperature is found to more often affect ecotype occurrence probabilities than origination probabilities. In most cases, there is a negative association between temperature and probability of occurring or first originating; this means that if temperature decreases, we would then expect an increase in the probability of occurring or first originating. In contrast, temperature range is estimated to be a good predictor of survival in only to mammal ecotypes and only marginally for one of those (Table \ref{tab:surv_temp}). Additionally, both of these cases have positive relationships, meaning that if temperature decreases it is expected that species survival will also decrease. 

The result that temperature does not affect the survival probability of most ecotypes is consistent with previous analysis of mammal diversity \citep{Alroy2000g}. The result that temperature affects origination probability, on the other hand, is in strong contrast to the results \citet{Alroy2000g}. An important difference between the anlayses presented here and that of \citet{Alroy2000g} is I am considering the effect of temperature on the probability of a species originating, assuming it hasn't originated yet while \citet{Alroy2000g} analyzes the correlation between the first differences of the origination and extinction rates with an oxygen isotope curve \citep{Zachos2001}. Origination or extinction rates have very different properties than the origination probabilities estimated here brought upon by the difference both in definition and units. Origination probability is the expected probability that a species that has never been present and is not present at time \(t\) will be present at time \(t + 1\); origination probability is defined for a single species. In contrast, per capita rates are defined (for origination) as the expected number of new species to have originated between time \(t\) and \(t + 1\) given the total number of species present at time \(t\); per capita rates are defined for the standing diversity. It is also important to note that even though there is an edge effect at the last time interval that causes an increase in the occurrence and origination probabilities of some ecotypes (Fig. \ref{fig:eco_occur}, \ref{fig:eco_origin}, the corresponding rates and population level birth/death dynamics do not share that pattern (Fig. \ref{fig:macro_values}, \ref{fig:ecotype_birth}, \ref{fig:ecotype_death}). However, it is still possible that the finding that temperature has an effect on origination may simply be because as time approaches the present the number of species which have originated increases and not because of climatic forcing of origination. 

All environmental factors are found to affect the occurrence and origination probabilities for most, but not all, mammal ecotypes (Fig. \ref{fig:group_pure_presence}, \ref{fig:group_origin_bd}). In contrast, the environmental factors probably do not affect differences in ecotype survival probability (Fig. \ref{fig:group_surv_bd}). The focus in previous research on temperature and major climatic or geological events without other measures of environmental context may have led to confusion in discussions of how the ``environment'' affects mammal diversity and diversification \citep{Alroy2000g,Figueirido2012}. The environment or climate are more than just global or regional temperature, it is also the set of all possible biotic and abiotic interactions that can be experienced by a member of the species pool. By including more descriptors of species' environmental context than simple an estimate of global temperature a more complete ``picture'' of the diversification process is inferred.

Analysis of relationship between temperature and origination rate is probably better suited for a continuous-time birth-death or multilevel stochastic differential equation model instead of a discrete-time model because the both continuous models estimate rates while discrete time models estimate probabilities \citep{Allen2011}. The \texttt{PyRate} model(s) are based on a continuous-time birth-death process \citep{Silvestro2014a,Silvestro2015b}. Unfortunately, a continuous-time model may be unsuited for most paleontological data as the fossil record is naturally discrete; fossils are assigned to temporal units, such as stages, which have age ranges. Individual fossils are not assigned individual numeric ages. This reality was in fact my one of motivations for using discrete-time birth-death model instead of one in continuous-time. There are of course exceptions to this characterization; the fossil record of graptolites from the Ordovician and Silurian \citep{Crampton2016a} and the fossil record of some mammal orders from Neogene are of high enough resolution that the application of continuous-time models is appropriate and less fraught.

The effect of species mass on either occurrence or origination and extinction was not allowed to vary by ecotype or environmental context. The primary reason for this modeling choice was that this study focuses on ecotypic based differences in either occurrence, or origination and extinction. Allowing the effect of body size to vary by ecotype, time, and environmental factors would increase the overall complexity of the model beyond the scope of the study. Instead, body size was included in order to control for its possible underlying effects \citep{McElreath2016}. A control means that if there is variation due to body mass, having a term to ``absorb'' that effect is better than ignoring it. %Additionally, the effect of body size was allowed to have a second-order polynomial form and no higher order polynomials were considered; this was done because it is hard to conceive of a more complex third- or higher-order relationship between body size and the other parameters. Finally, parametric forms of nonlinearity have not previously been considered, so the simple act of estimating a potential second-order relationship is an opportunity to test more complex hypotheses of the relationship between body size and both macroevolutionary and macroecological processes.

The only covariate allowed to affect sampling probability was mass and only as a linear predictor. Other covariates, such as the environmental factors considered here, could have affected the underlying preservation process that limits sampling probability; their exclusion as covariates of sampling/observation was the product of a few key decisions: model complexity, model interpretability, the scope of this study, and a lack of good hypotheses related to these covariates to warrant their inclusion. %It should be noted that in other similar studies that use a hidden Markov model, like both models in this study, to handle simultaneous estimation of sampling, origination, and extinction have not considered the possible effects of covariates, both species traits and environmental factors, on sampling CITATION.

%The time scale available with paleontological data is much greater than that obtainable from modern ecological studies, even long running observations CITATION. Specifically, the temporal scale of paleontological data allows for the complete turnover of a species pool to be observed, something that is impossible in ``real time.'' However, paleontological data is very limited in its spatial resolution, so the analysis of how the ecotypic diversity local communities change over time and how that is also the product of larger scale regional turnover remains unanswered. Phylogenetic comparative community ecology and phylogenetic comparative biogeography also discusses how the macroevolutionary processes helps structure an observed community, though it is not necessarily phrased that way. However, that community did not form in isolation but it the result of many factors interacting over time including incumbency, competition, limiting similarity, etc.

%The potential effects of common ancestry (i.e. phylogeny) on origination and extinction are not directly considered in this analysis. While a birth-death process approximates the speciation-extinction process underlying the phylogeny \citep{Silvestro2014a} this is not same as considering how the similarity between closely related species may affect the estimates of the effects of species traits to environmental factors on both origination and extinction \citep{Smits2015b,Harnik2014}. The inclusion of phylogeny can help disentangle if the functional composition of species diversity is shaped either by many closely related species occurring at the same time or if closely related species are more evenly distributed in time; this is analogous to if species within local communities are clumped or dispersed relative to their relatedness \citep{Webb2002,Kraft2007a,Cavender-Bares2009}. One of the principal barriers to the inclusion of the effect of phylogeny in either the pure-presence or birth-death models is computational; with well over 1000 tips, the calculation of the scale parameter defining the phylogenetic effect would be very slow and further increase the already slow computation time necessary for the marginalization of all possible discrete occurrence histories for \(z\).


\subsection*{Conclusions}

These results add a considerable degree of nuance to the narrative of changes to North American diversity being gradual. My results support the conclusions that ecotypic diversity is shaped more by changes to origination than extinction and that major changes to total diversification rate can be attributed to increases in origination of only some ecotypes. There are a number of interesting estimated ecotype diversity patterns. While arboreal ecotypes are diverse in the Paleogene, by the Neogene all arboreal ecotypes dramatically decreased in diversity and became either rare or absent from the regional species pool. The other ecotypes that decrease in diversity over the Cenozoic are plantigrade and scansorial insectivores and scansorial omnivores. The only ecotypes that demonstrate a sustained pattern of increasing diversity are digitigrade and unguligrade herbivores. When the environmental covariates analyzed here are inferred to affect the diversification of an ecotype, this effect is virtually always on origination and not survival. This analysis provides a much more complete picture of North American mammal diversity and diversification, specifically the dynamics of the ecotypic composition of that diversity. By increasing the complexity of analysis while precisely translating research questions into a statistical model, the context of the results is much better understood. Future studies of diversity and diversification should incorporate as much information as possible into their analyses in order to better understand or at least contextualize the complex processes underlying that diversity.





\end{document}
