\documentclass{beamer} 
\usepackage{amsmath,amsthm}
\usepackage{graphicx,microtype,parskip}
\usepackage{caption,subcaption,multirow}
\usepackage{attrib}

\frenchspacing

\usetheme{default}
\usecolortheme{whale}

\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]

\setbeamercolor{title}{fg=blue,bg=white}

\setbeamercolor{block title}{fg=white,bg=gray}
\setbeamercolor{block body}{fg=black,bg=lightgray}

\setbeamercolor{block title alerted}{fg=white,bg=darkgray}
\setbeamercolor{block body alerted}{fg=black,bg=lightgray}


\AtBeginSection[]
{
  \begin{frame}
    \tableofcontents[currentsection]
  \end{frame}
}

\title{Fall 2016 committee meeting for Peter Smits}
\author{}
\institute{}
\date{}


\begin{document}

\begin{frame}
  \maketitle
\end{frame}

\begin{frame}
  \tableofcontents
\end{frame}

\section{Post-doc applications}
\begin{frame}
  \frametitle{What I've applied for}
  \begin{itemize}
    \item project idea: models of both within- and across-species continuous trait evolution
    \item Miller Fellowship at Berkeley with Charles Marshall
    \item Killam Fellowship at UBC with Matt Pennell
  \end{itemize}
\end{frame}

\section{Chapter 2: brachiopod survival}
\begin{frame}
  \frametitle{Revisions following rejection}

  \begin{itemize}
    \item reviews considered text to be both too ``paleontological'' and technical
    \item new target is \textit{American Naturalist}.
    \item conceptual figures (e.g. behaviour Wiebull distribution, model structure)
    \item text emphasis on fitness, selection interpretations
  \end{itemize}
\end{frame}

\section{Chapter 3: mammal species pool}
\begin{frame}
  \begin{block}{What I've got done already.}
    \begin{itemize}
      \item model developed
        \begin{itemize}
          \item poor fit as is, species do not persist long enough
          \item TODO: preservation as function of body size
        \end{itemize}
      \item variational Bayesian inference
        \begin{itemize}
          \item proof of concept/approximate inference
          \item fit is still slow
          \item \textit{Folk Theory of Statistical Computing}?
        \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \begin{block}{What's left with this project?}
    \begin{itemize}
      \item full posterior inference
        \begin{itemize}
          \item extremely slow b/c latent variables
          \item \textit{Folk Theory of Statistical Computing}?
        \end{itemize}
      \item improve model structure
        \begin{itemize}
          \item preservation as function of body size
          \item second-order Markov process?
        \end{itemize}
      \item write manuscript
        \begin{itemize}
          \item no strong conclusions yet
          \item journal target?
        \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\section{Finishing}
\begin{frame}
  \begin{huge}
    See ``Timeline''
  \end{huge}
\end{frame}

\begin{frame}
  \frametitle{Options}
  \begin{itemize}
    \item Finish in the Spring or finish in the Summer.
    \item Idea
      \begin{itemize}
        \item Turn in, defend at end of Spring quarter.
        \item Technically finish Summer quarter.
        \item Unsure of details; need to contact Chair/Carolyn.
      \end{itemize}
  \end{itemize}
\end{frame}

\end{document}
