\documentclass[11pt,letterpaper]{article}

\usepackage{amsmath, amsthm}
\usepackage{graphicx,hyperref}
\usepackage{microtype, parskip}
\usepackage{bibentry}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{docmute}
\usepackage{lineno}

\frenchspacing

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{plain}                                                      %%
%%%%%%%%%% EXACT 1in MARGINS %%%%%%%                                   %%
\setlength{\textwidth}{6.5in}     %%                                   %%
\setlength{\oddsidemargin}{0in}   %% (It is recommended that you       %%
\setlength{\evensidemargin}{0in}  %%  not change these parameters,     %%
\setlength{\textheight}{8.5in}    %%  at the risk of having your       %%
\setlength{\topmargin}{0in}       %%  proposal dismissed on the basis  %%
\setlength{\headheight}{0in}      %%  of incorrect formatting!!!)      %%
\setlength{\headsep}{0in}         %%                                   %%
\setlength{\footskip}{.5in}       %%                                   %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                                   %%
\newcommand{\required}[1]{\section*{\hfil #1\hfil}}                    %%
\renewcommand{\refname}{\hfil References Cited\hfil}                   %%
\bibliographystyle{abbrvnat}                                           %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\setcounter{secnumdepth}{0}
\section{Budget justfication}

Funding is requested over two years: \textbf{\$4060.00} in Year 1 and \textbf{\$2500.00} in Year 2 in direct costs.

A) Senior Personnel: \$0

B) Other Personnel: \$0

C) Fringe Benefits: \$0

D) Permanent Equipment: \$0

E) Travel:

\begin{tabular}[H]{p{7.5cm} p{4.5cm} r}
  Budget Item & Item Cost per Unit & Item Total \\
  \hline
  Airfare: Chicago, IL, USA to Buenos Aires, Argentina (round trip) for co-PI & \$1500.00 per round trip (Year 1) & \$1500.00 \\
  Lodging: Buenos Aires, Argentina & \$80.00/night x 8 nights (Year 1) & \$640.00 \\
  Transport: Buenos Aires, Argentina & \$10.00/trip x 2 trips/day x 7 days (Year 1) & \$140.00 \\
  Transport: Buenos Aires, Argentina to La Plata, Argentina (round trip) for co-PI (Year 1) & \$10.00/trips x 2 trips & \$20.00 \\
  Lodging: La Plata, Argentina & \$100.00/night x 6 (Year 1) & \$600.00 \\
  Airfare: Chicago, IL, USA to Austin, Texas, USA (round trip) for co-PI for 2016 Evolution Meeting & \$500.00 (Year 2) & \$500.00 \\
  Lodging: Austin, Texas, USA for co-PI for 2016 Evolution Meeting & \$120.00/night x 5 nights (Year 2) & \$600.00 \\
  Airfare: Chicago, IL, USA to Seattle, Washigton, USA (round trip) for co-PI for 2016 Geological Society of America Meeting & \$600.00 (Year 2) & \$600.00 \\
  Lodging: Austin, Texas, USA for co-PI for 2016 Geological Society of America Meeting & \$120.00/night x 5 nights (Year 2) & \$600.00 \\
  Transport: Miscellaneous travel to/from Chicago, IL airport for co-PI & \$50.00/trip x 2 directions x 4 trips (Year 1 + 2) & \$400.00 \\
  & \textbf{Total} & \$5600.00 \\
  & \textbf{Total} & \$5600.00 \\
  & \textbf{Requested from NSF} & \$5600.00
\end{tabular}

This DDIG research involves traveling to museums in Argentina and measuring specimens. I (PDS) will then spend two weeks in the vertebrate paleontology collections at both the Museo Argentino de Ciencas Naturales in Buneos Aires and Museo de La Plata in nearby La Plata. Both of these museums house large collections of fossil Cenozoic mammals from which I do not have body mass estimates.

I will fly into Buenos Aires, Argentina. From there, I will spend one week in the city while visiting the Museum Argentino de Ciencas Naturales. The second week I will travel by bus to the nearby city of La Plata. While staying in La Plata, I will be visiting the collections at the Museo de La Plata. 

In the second year, I am requesting funds to present some of my results at the Evolution Meeting held in Austin, Texas in June 2016. I am also requesting funds to present more of my results at the Geological Society of America meeting in Seattle, Washington in October 2016.

All flight and hotel cost estimates were obtained through Kayak.com. Transportion cost estimates were obtained through TripAdvisor.com.


F) Participant Support Costs: \$0

G) Other Direct Costs:

1) Materials and Supplies

\begin{tabular}[H]{l r}
  Budget Item & Item Total \\
  \hline
  Mitutoyo 12\" Calipers 500-196-20 & \$160.00 \\
  4TB external harddrive & \$200.00 \\
  \textbf{Total} & \$360.00 \\
  \textbf{Requested from NSF} & \$360.00
\end{tabular}

Mitutoyo calipers are standard tools for obtaining exact measurements and are essential for my data collection. Currently, I am using a pair on departmental loan however these are of limited access because they are shared between many students. 

External harddrive is for improving the backup, security, and preservation of all data collected and used in this analysis.




2) Consultant Services

\begin{tabular}[H]{l l r}
  Budget Item & Item Cost per Unit & Item Total \\
  \hline
  Spanish Language Tutor and Consultant & \$70.00/session x 10 sessions & \$700.00 \\
  & \textbf{Total} & \$700.00 \\
  & \textbf{Requested from NSF} & \$700.00
\end{tabular}

I cannot speak Spanish though I am capable of understating technical papers due to fluency in French. One of my major goals is to attempt to translate and transmit information ``locked'' in Spanish-language journals, so I need tutoring in Spanish along with translation consultation. Training in Spanish will greatly aide communication with experts in South America and possibly lead to presentation and outreach opportunities.


H) Total Direct Costs: \$6560

I) Total Indirect Costs: \$3805

Indirect cost rate is 58\% (of Modified Direct Cost = Total Cost less than Tuition) for the entire project period, per the indirect cost rate agreement dataed 01/18/2013 between the University of Chicago and its cognizant federal agency, the National Science Foundation.


J) Total Direct and Indirect Costs: \$10365

K) Residiual Funds: \$0

L) Amount of this Request: \$10365.00


\end{document}
