\documentclass{article}

\usepackage{amsmath, amsthm}
\usepackage{setspace}
\usepackage{microtype, parskip}
\usepackage[comma,sort&compress]{natbib}
\usepackage{lineno}
\usepackage{docmute}
\usepackage{caption, subcaption, multirow, morefloats, rotating}
\usepackage{wrapfig}

\frenchspacing

\title{Timeline 2016 forward}
\date{}

\begin{document}
\linenumbers
\modulolinenumbers[2]

\maketitle

\section{Winter quarter 2016}
Next steps for current projects?
\begin{itemize}
  \item Brachiopod regional diversification
  \item Mammal trait occurrence ratio
\end{itemize}


\section{Spring quarter 2016}
TAing
\begin{itemize}
  \item Hopefully Dimitry Kondrashov's Biological Modeling class.
  \item Otherwise try to snag something in Greg and Stefano's new E\&E class
\end{itemize}

Evolution (June 17-21 Austin Texas)
\begin{itemize}
  \item Present on which project?
    \begin{itemize}
      \item Brachiopod diversity is a more mature project.
    \end{itemize}
  \item Suggestions for people to talk about post-doc opportunities?
\end{itemize}

Committee meeting
\begin{itemize}
  \item What are the goals between now and then?
  \item What should I have written (and sent to MORE people than just Ken or Michael)?
\end{itemize}


\section{Summer quarter 2016}
Finishing and post-doc applications
\begin{itemize}
  \item Am I finishing in 2017? 
  \item Do I plan for a 6th year?
  \item Post-doc where and with whom?
\end{itemize}

Project goals
\begin{itemize}
  \item Try to have one of the current projects in late editing phase?
\end{itemize}


\section{Fall quarter 2016}
Many post-doc applications due
\begin{itemize}
  \item Miller Fellowship w/ Charles Marshall? I've not spoken to him about this yet.
\end{itemize}

GSA September 25-28 Denver Colorado
\begin{itemize}
  \item Present on which project?
\end{itemize}

SVP October 26-29 Salt Lace City Utah
\begin{itemize}
  \item Do I need to go?
  \item What would I present? Romer talk?
\end{itemize}

Committee meeting
\begin{itemize}
  \item What are the goals between now and then?
\end{itemize}


\section{Winter wuater 2017 and beyond}
Finishing my dissertation
\begin{itemize}
  \item What projects are ``in'' as chapters?
  \item What do I need to do to finish?
    \begin{itemize}
      \item Dissertation office prep starts 6 months prior.
    \end{itemize}
\end{itemize}


\end{document}
