\documentclass{article}

\usepackage{amsmath, amsthm}
\usepackage{setspace}
\usepackage{microtype, parskip}
\usepackage[comma,sort&compress]{natbib}
\usepackage{lineno}
\usepackage{docmute}
\usepackage{caption, subcaption, multirow, morefloats, rotating}
\usepackage{wrapfig}

\frenchspacing

\title{Peter Smits Timeline Spring 2016 forward}
\date{}

\begin{document}
\linenumbers
\modulolinenumbers[2]

\maketitle

\section{Spring 2016 (current quarter)}
TAing 
\begin{itemize}
  \item Kondrashov's Quantitative Modeling for Biology.
  \item The kind of class I'll probably be asked to teach.
\end{itemize}

Evolution 2016
\begin{itemize}
  \item Brachiopod survival project
    \begin{itemize}
      \item submitted to Evolution
      \item complete and ``fancy''
    \end{itemize}
  \item Opportunity to build my ``brand.''
\end{itemize}


\section{Summer 2016}
Post-doc applications
\begin{itemize}
  \item Make contact with all prospective post-doc advisors (see presentation).
  \item Write research statement.
    \begin{itemize}
      \item Need a review of my ``research program''.
      \item Second self-generating research line.
    \end{itemize}
\end{itemize}


\section{Fall 2016}
Post-doc applications
\begin{itemize}
  \item Miller Fellowship 12/10 (nomination), 1/10 (invited application)
  \item Smithsonian Fellowship 12/1
  \item NIMBioS Fellowship 9/1 or 12/11
  \item Michigan Fellowship
\end{itemize}

Conferences
\begin{itemize}
  \item mammal regional species project
  \item GSA
  \item SVP
    \begin{itemize}
      \item Romer Talk?
    \end{itemize}
\end{itemize}

Teaching
\begin{itemize}
  \item I might be TAing the BSD entering first-year stats/programming boot camp
  \item I've agreed to teach the TA training class again
\end{itemize}


\section{Winter 2017}
Dissertation
\begin{itemize}
  \item Meet with Dissertation Office to go over formatting. 
    \begin{itemize}
      \item I have a \LaTeX template already.
    \end{itemize}
  \item Apply to graduate in Spring quarter? Must decide by first week of that quarter.
  \item Pretty much has to be completely written by this point.
\end{itemize}


\section{Spring 2017} 
(assumes I am finishing spring/summer 2017)

Dissertation
\begin{itemize}
  \item Process for finishing.
    \begin{itemize}
      \item Get draft of dissertation to entire committee by end of first week of final quarter.
      \item Need to email Chair to request approval of committee, advisors, and defense date.
      \item Defense has to take place around 2nd week of quarter.
      \item Need to submit final dissertation by 5th week of quarter to graduate that quarter.
    \end{itemize}
  \item Possibility of defending in Spring but graduating in Summer.
    \begin{itemize}
      \item More time for revisions.
      \item Get paid for one more quarter.
    \end{itemize}
\end{itemize}

Evolution 2017
\begin{itemize}
  \item mammal regional species project
  \item Ernst Myer talk? (Systematic Biology)
  \item W D Hamilton talk? (Evolution)
  \item I probably wouldn't win either of these.
\end{itemize}


\section{Summer 2017}
Dissertation?
\begin{itemize}
  \item If not in spring, then summer.
  \item All of the above still applies.
\end{itemize}



\end{document}
